### 2021-08-15 papers/source/rewriting/pldi-2020-reviews-response-rejection.txt

The response to this is hinted at on [1120]-[1123]: "As the
reification method was not especially optimized, and there exist fast
reification methods [13]".  However, refication is not actually the
bottleneck.  The bottleneck is the final cbv at the end, and this is
only due to a bug in Coq's implementation of NbE reduction that makes
substitution quadratic in the number of
binders (we wish we could link to the bug tracker, but it might reveal
something about our identities).
While we felt it would be misleading to give only the "only rewriting"
numbers, we do believe this is really the benchmark that matters, Coq
bugs aside.  (And we hope this bug in Coq will be fixed soon.)

Responses to Review C
=====================

Thank you for your points; we will aim to incorporate your feedback in
the next revision of the paper.

One clarification we'd like to make is that the abstract interpreter
itself is beyond the scope of this project; we merely want to
demonstrate that it is possible to integrate separate abstract
interpretation passes with the partial evaluator and rewriting
framework.

Responses to Review D
=====================

> Why not implement, once and for all in Coq, the efficient rewriting
> engine algorithms, based on indexing and pattern match automata,
> that the term rewriting community has developed over the years and
> implemented in truly fast rewrite engines, performing millions, or
> even tens of millions, or rewrite steps per second?  That would
> benefit the entire Coq community.

Indeed, this would be a great next step.  We started with a somewhat
more modest goal, and already found the proof effort significant.

> Why should one trust the rewrite rules as correct? Don't they become
> part of the trust base anyway?

To be accepted by the engine, the rewrite rules must be proven in Coq.
So no unsound rewrites are permitted (unless you use unsound axioms to
prove them). The checking happens as part of the step from line 502 of
the paper, where we bring all the rewrite rules together in a list of
dependently typed records, each one pairing the syntax of a rewrite
rule with a required proof of its soundness, stated in terms of
interpretation of the reified syntax tree.

> Also, isn't your term translation back and forth to Coq the same as
> like giving a semantics to an ML-like language like in [1]?  One
> still has to trust these translations, including the fact that the
> ML-like language and Coq share the semantics of their similar
> constructs., which to me counts as the same level of trust that [1]
> requires from their users.  You should clarify that in the paper.

The translations in our tool are *not* trusted. This is because every
time our code reifies a term, the Coq kernel checks that the embedded
AST our code produced has the same semantics as the original term by
unfolding the interpreter applied to the AST and observing that it
results in the original term. Our rewriting transformation is proven to
be semantics-preserving, so the theorem we get at the end is that the
new term, post-rewriting, has the same semantics as the original term.
Since the kernel checks the correctness of the semantics of the initial
term, there's nothing additional here to trust. We should perhaps
clarify this in the paper.

> Lines 701-704: this becomes part of the trust base, no?

Thankfully, it does not.  In this case, reify and reflect are internal
operations which are not exposed in the end-to-end theorem (any moreso
than code for CompCert's intermediate phases appears in a generated
theorem witnessing correct compilation of a concrete C program), and
we prove the properties that we need of these operations.

> How is the approach in [11] different from the approach here?

> Is it only a performance improvement?

[11] relies on a delicate ad-hoc combination of hand-writing all of
the code in CPS to take advantage of built-in reduction for partial
evaluation, and then using a hand-rolled domain-specific rewriter to do
the necessary algebraic simplifications. The authors of [11] describe
the need to write all algorithms in CPS as "arbitrary and unfortunate."
Further, this two-stage approach requires reifying large blocks of code,
and reifying repeatedly for every prime, which was prohibitively slow in
many cases ([11] Appendix B).

The approach described here lets us reify the generic code once, and
combine all of the rewriting and partial evaluation into the same
generic framework. We believe this is much easier to use. Probably the
clearest example of this is that our rewriting and partial evaluation
can be applied to code that was written without considering our tool as
a target, whereas the code repository accompanying [11] contains two
versions of many functions: one human-readable, and one for partial
evaluation (which they prove equivalent each time).

The performance improvement is also essential, in the sense that some of
the primes that did not finish even after waiting nearly a thousand
hours in [11] finished for us in seconds, minutes, or, at most, an hour
or two.

> How about the trust-base?

Running our rewriter in Coq has the same trusted base as [11].  The
faster version of our code also relies on extraction to OCaml, but this
is not essential.

> I was a bit disappointed to read that you generate essentially the
> same code as [11] and you only care about the performance of the
> partial evaluator; it would have been a really nice application
> where correctness is guaranteed through Coq, but if the application
> already had all these guarantees, then the work in this paper feels
> less critical and more incremental.

Again, we would like to emphasize that many primes could not be
handled at all by [11] even with hundreds of hours and 64 GB of RAM.
The "performance improvement" here allows us to not just handle the
same primes faster, but also allows us to handle primes that could not
be handled at all.

### 2021-08-15 _site/papers/2015-parsing-parse-trees-reviews/detailed-review-response-notes.txt

Reviewer A

> Splitting oracles: the essential algorithmic component of parsing is
  the association of the input into meaningful trees. Deferring the
  split choice to an "oracle" means that all non-trivial computation
  is in it. So I would not say that you're performing parsing.

This is a good point.  We are intentionally not doing anything interesting with regard to algorithmic parsing theory; that is reserved for future papers on currently in-progress work on synthesizing efficient parsers.  We picked the simplest, most obvious parsing algorithm we could think of, as a starting point.  That you claim our algorithm contains only trivial computation speaks to our success in picking a simple starting point.  Allow me to add a trivial splitter, which allows us to really perform parsing, trivially: the brute-force splitter, which returns the sequence of numbers [0, 1, ..., length of string] is sufficient for correctness, because we filter the list of splits to accept only valid ones.  We believe that the idea of parsing parse trees is complicated enough as it stands, and does not need extra complexity in exposition introduced by chosing a non-trivial parsing algorithm as a starting point.

> The work that remains appear to be mostly about detection of
  recursion in the grammar, which is easier than the above step.

Certainly not easier, when good performance is not demanded (as it is not, for code that is never run and only compiled).  It is straightforward, bordering on trivial, to prove that returning every possible split of the string does not miss any splits (and is thus complete), and it does not add any complexity to the completeness proof to do it this way.

> The textbook approach to detect loops is by reduction to Chomsky
  Normal Form. (This can be done statically on the grammar.) In
  comparison your approach feels awkward. At the minimum, discuss
  your choice!

This seems more complicated, but I am not sure.  Computing the Chompsky Normal Form of a grammar in a manifestly terminating way, proving that every grammar has a Chompsky Normal Form, that a grammar has a parse iff its CNF has a parse, this seems more complicated than proving that our way of keeping track of which nonterminals have been seen does not lose us any valid parses.  Perhaps it is not more complicated, but it seems less ovbious, and not obviously simpler nor more straightforward.

> I don't think that your interface to the parser is sensible. Why is
  it so long (fig 1)? In my mind I should just supply the grammar
  rules. (And perhaps your oracle.) I can also not reconcile your
  explanation (reading it off the minimal parse tree) with fig 1.

The interface is not a user-facing one; it is the internal one, to be partially instantiated to generate minimal parse trees, or parse trees, or booleans.  Once you fix the type of the return, indeed, all that is left is to specify the rules of the grammar, and, perhaps, the splitter.  The length of the interface results from the fact that the algorithm, a priori, knows nothing about minimal parse trees; the interface needs to have enough generality to specify exactly how to build a minimal parse tree, and how to prove that no such parse tree exists.  Hence it has enough cases for all the constructors, and enough cases to prove absurdity from impossible parse trees.

>   1. Define a "ParseTreeOf nt s" (the spec)
  2. Write an algorithm to decide the relation:

     parse : (nt : NT) -> (s : String) -> Decide (MinParseTreeOf ns t)

  And to get to the ParseTreeOf version only the de-loop lemma is
  needed (in one direction).

The de-loop lemma would need to be strengthened to do this; the current version only de-loops at one level, which is not enough to get ParseTreeOf -> MinParseTreeOf.  When proving correctness of the parser returning booleans directly, the full delooping construction is longer than the entire rest of the soundness and completeness proofs, combined!  Note that the full delooping construction includes the generation of ParseTreeOf from MinParseTreeOf as a way of getting well-foundedness evidence.

Reviewer B
> I had problems in matching the paper to the accompanying Coq codes.

I agree, this is a problem.  The accompanying code is a half-assed proof-of-concept, not an elegant match for the description.

I agree, this is a problem.  We did not find enough time to clean up the code.  It was an explicit goal in writing the paper to hide as many of the warts of our code-base as we could: long names, long lines, costly mistakes that we didn't have time to fix, unenlightening details forced on us by Coq.

> E.g., grep could not find has_parse or min_parse -- I was interested in a proof of min_parse.)

We apologize.  [min_parse] is named [minimal_parse_nonterminal__of__parse] in our code base.  We should have done more to make the code match the paper, and intend to do so, should the paper be accepted.
[FIXME: Adam: Is this okay to promise?]

> I miss technical comparisons
with the work by Ridge in HOL4. The authors say their approach is different from Ridge's in that they use dependent types which have a different aesthetic appeal 
-- the statement is too general to be informative.

FIXME: How to respond?

Reviewer C:
> what I was wanting to see was a set of versions of increasing complexity (with diffs indicated) that gave more of an indication of the way in which the development had proceeded.

To a first approximation, there is the code in 1.1, the slightly more complicated code in 1.2, and the fuller code of Figure 2.  Perhaps we could have indicated diffs more clearly, or rewritten 1.1 and 1.2 when it came time to introduce Figure 2, perhaps indicating which bits were already there in 1.1, which bits where there to prevent infinite regress from 1.2, and which bits are just there to handle the more complicated type signature.  In the actual code, there is BooleanRecognizer.v, which corresponds to the grammar-independent version of the code in 1.2 (equivalently, to the code in Figure 2 specialized to booleans), while DependentlyTyped.v gives the full code of the interface and implementation of Figures 1 and 2.

> there is a substantial amount to do on defining splitters and their effect on performance (more discussion on splitters below); this is reported as planned work;

Yes, this paper is about parsers that can prove their own correctness.  We should have mentioned that we believe this can be extended to other parsing algorithms.

> when splitting strings, this is not the case and it was not clear (at least to me) that for every grammar a splitter of the right kind exists

The brute-force splitter, which returns the sequence of numbers [0, 1, ..., length of string] is sufficient for correctness, because we filter the list of splits to accept only valid ones.

> did editors and any other tools support the development, or not?

We used ProofGeneral, with it's interactive display of coqtop feedback, syntax highlighting, and saving capabilities.  But most of these features are built into coqtop, and are assumed when developing Coq code; an interactive compiler seems essential when developing proof scripts and dependently typed code.  Is it worth mentioning explicitly?

Reviewer D

> The biggest criticism is about the code. The presented code in the
paper is not intact with the real development which makes it
impossible to really see if the proposed approach is as simple and
elegant as authors claim. I was unable to find functions like
"has_parse", "has_no_parse", "min_parse", "parse_full", 
"parse_sound", "parse_complete", etc.

Yes, this needs to be fixed, thank you for the suggestions.

> You are using provocative statement that parser is proving its
correctness and this is not what is happening.

I am tempted to try to do this precisely, rather than approximately, as I suggest in the last section.

> At page 2 you say that your strategy works only with finite sets of
nonterminals.  However, set of nonterminals could be infinite, but
then you should add the step of extracting the "used" nonterminals
from the grammar.

Consider A := B, B := C, C := D, and so on, ad infinitum.  There is no finite set of "used" nonterminals to be extracted, not even a decision strategy to see if it terminates.  The strategy really only works for manifestly finite sets of explicitly given "used nonterminals", which is what is meant by "finite sets of nonterminals".

> In subsection 2.3 it would be nice if you would make precise which
parse tree is returned by the parser for ambiguous grammars. In that
sense your notion of "completeness" is very special.

It depends on the splitter.  It returns the first loopless parse tree that the splitter provides it.  It would be relatively easy to make it return all parse trees, roughly by replacing a "fold" with a "flat_map".
